# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
library(dplyr)
library(ggplot2)

file <- 'pa1_activity.zip'
# download.file('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip',
#               method='curl', destfile=file)
# unzip(file)

activity <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?
```{r}
dailySteps <- group_by(activity, date) %.% summarise(steps=sum(steps, na.rm=T))
ggplot(dailySteps, aes(steps)) + geom_histogram()
mean(dailySteps$steps)
```

Mean daily steps = `r mean(dailySteps$steps)`

## What is the average daily activity pattern?
```{r}
intervalSteps <- group_by(activity, interval) %.% summarise(steps=mean(steps, na.rm=T))
intervalSteps
ggplot(intervalSteps, aes(x=interval, y=steps)) + geom_line()
mostActiveInterval <- intervalSteps[which.max(intervalSteps$steps), "interval"]
mostActiveInterval
```

## Imputing missing values
I am using the mean number of steps during a particular time interval over all the days in the place of the missing value.

```{r}
numMissing <- sum(is.na(activity$steps))
numMissing

missingInds <- which(is.na(activity$steps))
intervalMissing <- data.frame(interval=activity[missingInds, "interval"])
activityNew <- activity
activityNew[missingInds, "steps"] <- (left_join(intervalMissing, intervalSteps, by="interval"))$steps

dailyStepsNew <- group_by(activityNew, date) %.% summarise(steps=sum(steps, na.rm=T))
ggplot(dailyStepsNew, aes(steps)) + geom_histogram()
mean(dailyStepsNew$steps)
median(dailyStepsNew$steps)
```

The mean number of daily steps after imputing missing values is `r mean(dailyStepsNew$steps)`. This is greater than before.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekday <- weekdays(as.POSIXct(activityNew$date))
activityNew$dayType <- rep("Weekday", nrow(activityNew))
activityNew$dayType[which(weekday %in% c("Sunday", "Saturday"))] <- "Weekend"
activityNew$dayType <- as.factor(activityNew$dayType)

tmp <- group_by(activityNew, interval, dayType)
intervalStepsNew <- group_by(activityNew, interval, dayType) %.% summarise(steps=mean(steps, na.rm=T))

ggplot(intervalStepsNew, aes(interval, steps)) + geom_line() + facet_grid(dayType~.) +
  xlab('Number of steps')
```

As seen from the above charts, overall activity is higher on weekends, but the highest peak is in the morning on weekdays. On weekends, the activity is relatively more uniform throughout the day.